### 2019-05-09 Writeup/PERT_Paper_Draft2.tex
Fixed bugs in sub.jl

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\renewcommand{\textfraction}{0.0}
\newcommand{\dst}{\displaystyle}
\newcommand{\minx}{\mbox{\( \dst \min_{x \in X} \)}}
\newcommand{\Efx}{\mbox{\( \dst E f (x, \xi) \)}}
\newcommand{\Efxhat}{\mbox{\( \dst E f (\hat{x}, \xi) \)}}
\newcommand{\hxx}{\mbox{\( \hat{x} \)}}
\newcommand{\bpi}{\bar{\pi}}
\newcommand{\xx}{\mbox{\( x \)}}
\newcommand{\txxi}{\mbox{\(\xi\)}}
\newcommand{\var}{\mbox{var}}
\newcommand{\cF}{{\cal F}}
\newcommand{\cG}{{\cal G}}
\newcommand{\cN}{{\cal N}}
\newcommand{\cO}{{\cal O}}
\newcommand{\txi}{{\xi}}
\newcommand{\PP}{\mbox{\(SP\)}}
\newcommand{\PPn}{\mbox{\(SP_n\)}}
\newcommand{\PPnx}{\mbox{\(SP_{n_x}\)}}
\newcommand{\noi}{\noindent}
\renewcommand{\ss}{\smallskip}
\newcommand{\ms}{\medskip}
\newcommand{\bs}{\bigskip}
\newcommand{\st}{\mbox{s.t.}}
\newcommand{\wpo}{\mbox{wp1}}
\newcommand{\iid}{\mbox{i.i.d.\ }}
\newcommand{\vsmo}{\vspace*{-0.1in}}
\newcommand{\vsmt}{\vspace*{-0.2in}}
\newcommand{\vso}{\vspace*{0.1in}}
\newcommand{\vst}{\vspace*{0.2in}}
\newcommand{\mc}{\multicolumn}
\newcommand{\cP}{{\cal P}}
\newcommand{\underv}{\mbox{$\underbar{$v$}$}}
\allowdisplaybreaks

@@

	The management of complex projects through optimization has a rich history in operations research, beginning with the critical path method of \cite{kelley1961criticalpath} (see S{\"o}derlund \cite{soderlund2004building} for an overview). A project is a collection of activities \(I\)
	 %of some duration \(D_i, i \in I\)
	 , and there are precedence relationships, \(\mathcal{A} \subseteq I \times I\), between activities due to logical or technological considerations. A project can be characterized by an acyclic activity network \(\mathcal{G} = (I,\mathcal{A})\). An arc \((i,k) \in \mathcal{A}\) indicates that activity \(i\) has to be finished before the start of activity \(k\), and its length shows the required separation between their start time,
	 %duration of an activity 
	 \(D_{ik}\). An activity cannot start until all its predecessors are completed, while multiple activities can be processed at the same time without an upper limit, as long as the precedence requirements are satisfied. See Elmaghraby~\cite{Elmaghraby77} for a detailed treatment of activity networks. In this setting, ``crashing" is an action that consumes a certain amount of resource and shortens the duration of activities accordingly~\cite{kuhl2008dynamic}. We can apply a finite set of crashing options, \(j \in J_i\), to activity \(i \in I\). One unit application of each option incurs a cost of \(b_{ij}\), \textcolor{blue}{while it decreases all required separation lengths starting from activity \(i\) by \(D_{ik}e_{ij},\ \forall (i,k) \in \mathcal{A}\); i.e., the crashed duration is proportional to both the nominal separation length and the crashing decision.} The total cost of crashing cannot exceed a given budget, \(B\). The deterministic optimization problem for crashing an activity network was proposed in the 1960s \cite{fulkerson1961network, kelley1961criticalpath}. The goal is to complete all activities in minimum time by allocating resources under one or more budget constraints. We create two dummy activities \(S\) and \(T\) to represent the start and the end of the entire project. Activity \(S\) should precede every activity \(i \in I \backslash \{S\}\) and \(T\) should succeed every activity \(i \in I \backslash \{T\} \), and they both have zero duration, \(D_{Si} = D_{iT} = 0,\ \forall i \in I\). Given $t_S \equiv 0$, the objective is then to minimize the start time of activity \(T\). Thus, we formulate the deterministic project crashing problem as:
	% In constraint~\eqref{cons:dxub}, we can put an artificial upper bound \(1\) on all crashing decisions \(x_{ij}\) by taking proper values of \(e_{ij}\). \\

@@

	%The deterministic model can be extended to incorporate uncertainty of activity duration, starting with the program evaluation and review technique (PERT) \cite{malcolm1959application}. The PERT model assumes that activities are independently distributed with beta distributions, and the expected project span can be approximated by a normal distribution. However, this assumption may not be accurate for a general case~\cite{Elmaghraby77}. Therefore, assuming the activity duration follow a given probability distribution, Monte Carlo simulation methods can be used to estimate the expected project span, since it is is difficult to express the expected project span analytically \cite{burt1971conditional,van1963letter}. 

@@

	For a project crashing problem under stochastic disruptions, we assume at most one stochastic disruption can occur at a random time in the project span, since for the applications we are interested in (natural disasters, terrorist attacks, major market crashes, etc.), it is unlikely for two disruptions to occur in entire the time horizon. For example, suppose we manage a construction project, and we aim to plan against the potential hazard caused by earthquakes. It is unlikely for two major earthquakes to hit the same area twice within a year (aftershocks can be considered part of a major earthquake.) It is also reasonable to assume for each activity \(i \in I\), the crashing decision needs to be made at the start of that activity, and cannot be changed afterward, especially when contracts are involved in commitment of resources~\cite{oberlender1993project}. The disruption does not affect the started activities but changes the length of those that have not started according to a random distribution. It is usually hard to compute the recourse function directly when parameters are distributed according to continuous distributions, and therefore SAA is proposed to approximate the problem by a large-scale deterministic optimization problem with a finite scenario set \cite{kim2015guide,shapiro2009lectures}. In this paper, we assume there is a discrete finite set of scenarios indexed by \(\omega \in \Omega\). For each scenario \(\omega\), the random realization of parameters, \(\xi^\omega\), consists of the timing of disruption, \(H\), and the changed duration of each activity, \(d_i, \forall i \in I\). \\
	The notation for the model is displayed as follows:

@@

		\(D_{i}\)& \(\qquad\) & original duration of activity \(i\), \(i \in I\);&\\

@@

		\(B\) & \(\qquad\) & total budget;&\\
		\(b_{ij}\) & \(\qquad\) & cost of crashing option \(j\), \(i \in I, j \in J_i\);&\\

@@

	\noi The extensive formulation of the two-stage stochastic program is shown as formulation~\eqref{prob:extensive}:

@@

		\text{s.t.} \quad & t_k - t_i \geq D_{i}(1 - \sum_{j \in J_i} x_{ij} e_{ij}) \qquad \qquad \forall \,i \in I, (i,k) \in \mathcal{A} \label{cons:Sep}\\

@@

		& t_k^\omega - t_i^\omega \geq D_i + d_i^\omega G_i^\omega -\sum_{j \in J_i} D_i e_{ij} x_{ij}^\omega - \sum_{j \in J_i} d_i^\omega e_{ij} z_{ij}^\omega \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:scenSep}\\

@@

	In model~\eqref{prob:extensive}, we are minimizing the expected project span by taking the sum of each scenario's project span weighted by the scenario probability. We keep the constraints~\eqref{cons:dSep}-\eqref{cons:dSingleBudget} for the nominal scenario as~\eqref{cons:Sep}-\eqref{cons:SingleBudget}. In constraints~\eqref{cons:G1}-\eqref{cons:G2}, variables \(G^\omega_i\) takes value \(1\) if activity \(i\) starts after the disruption time; otherwise it takes value 0, and \(M\) is a large number to enforce this logic relationships. This is important in our problem setting because the duration of each activity depends on its temporal relationship to the disruption time, which is reflected in constraint~\eqref{cons:scenSep}. Also we have to ensure that the decisions made before the disruption time in each scenario are the same as the nominal decisions. Therefore, we set up non-anticipativity constraints~\eqref{cons:tG1}-\eqref{cons:xG2}. For each scenario, the duration of activity \(i\) becomes \((D_i + d_i^\omega G_i^\omega)(1 - \sum_{j \in J_i} e_{ij}x_{ij}^\omega)\), which expands to the form of constraint~\eqref{cons:scenSep}. If \(G_i^\omega = 0\), which means activity \(i\) starts before the disruption time of scenario \(\omega\), this expression is the same as \(D_i (1 - \sum_{j \in J_i} e_{ij}x_{ij})\) since \(x_{ij} = x_{ij}^\omega\) is enforced by constraints~\eqref{cons:xG1} and~\eqref{cons:xG2}. If \(G_i^\omega = 1\), then the duration of activity \(i\) is changed to \(D_i + d_i^\omega\). The expression \((D_i + d_i^\omega G_i^\omega)(1 - \sum_{j \in J_i} e_{ij}x_{ij}^\omega)\) contains a bilinear term \(G_i^\omega x_{ij}^\omega\), which can be linearized by introducing a binary variable \(z_{ij}^\omega\) and constraints \eqref{cons:linearize1}-\eqref{cons:linearize3}.

@@

	In this section, we prove that the project crashing optimization problem under a stochastic disruption is NP-hard. The proof process is largely based on the proof of De et al.~1997 for the discrete time-cost tradeoff (DTCT) problem. In De et al.~\cite{de1997complexity}, the authors show that an exactly-one-in-three 3SAT (EOIT\_3SAT) problem can be transformed to an instance of the DTCT decision problem in polynomial time. Since EOIT\_3SAT is NP-complete (see \cite{Garey1979ComputersAI} for the detailed proof), the authors prove that the DTCT decision problem is also NP-complete, and the DTCT optimization problem is NP-hard.\\

@@

	Similar to De et al.~\cite{de1997complexity}, we also first introduce the EOIT\_3SAT problem: let \(U = \{u_1,u_2, \dots, u_n\}\) be a set of variables. A literal can be either \(u\) or \(\bar{u} = \neg u\) for a \(u \in U\). Let \(C = \{c_1, c_2, \dots, c_m\}\) be a set of clauses, each of which is formed by a disjunction of three literals. The EOIT\_3SAT problem asks for whether there is a truth assignment for \(U\) such that each clause in \(C\) has exactly one true literal. \\

@@

	Based on the above instance of EOIT\_3SAT, we formulate an instance of the activity network for the project crashing problem. The activity network consists of three layers of nodes: the first layer contains \(4n\) nodes and represents the truth assignment of each variable; the second layer contains \(3m\) nodes and represents the value of clauses; the third layer contains node \(T\) which represents the end of the project. The first layer has \(n\) components in parallel, each of which corresponds to a variable and contains 4 nodes connected, as shown in Figure~\ref{fig:layer1}. Figure~\ref{fig:layer1} also displays the connection between the first layer and the third layer. We set the scenario set as a singleton \(\Omega = \{1\}\) and set the disruption time as \(H^1 = 0\) with probability \(p^1 = 1\). We set parameter values associated with the arcs in Figure~\ref{fig:layer1} as:

@@

	& D_{u_{j1},u_{j2}} = 0 \qquad \quad d^1_{u_{j1},u_{j2}} = 0  \\
	& D_{u_{j1},u_{j4}} = 2 \qquad \quad d^1_{u_{j1},u_{j4}} = -1  \\
	& D_{u_{j2},u_{j3}} = 1 \qquad \quad d^1_{u_{j2},u_{j3}} = 1\\
	& D_{u_{j4},T} = 0 \qquad \quad \;\; d^1_{u_{j4},T} = 0.

@@

	We do not allow any activities within the first layer to be crashed, i.e., \(J_{u_{jk}} = \emptyset\), for all \(j = 1,2,\dots, n,\ k = 1,2,3,4\). With this setup, the length of the critical path for the \(j\)-th component is always \(2\) for all \(j = 1,2,\dots,n\), as it is optimal to have activity \(u_{j1}\) start at time \(0\). However, whether the critical path traverses activities \(u_{j2}\) and \(u_{j3}\) (top path) or the activity \(u_{j4}\) (bottom path) depends on the value of \(G^1_{u_{j1}}\). If \(G^1_{u_{j1}} = 1\), the \(G\) variables will take value \(1\) for all subsequent nodes. The top path yields a length of \(D_{u_{j1},u_{j2}} + d^1_{u_{j1},u_{j2}} + D_{u_{j2},u_{j3}} +d^1_{u_{j2},u_{j3}} = 0 + 0 + 1 + 1 = 2\), while the bottom path yields a length of \(D_{u_{j1},u_{j4}} + d^1_{u_{j1},u_{j4}} = 2 - 1 = 1\), which makes the top path critical. If \(G^1_{u_{j1}} = 0\), it is optimal to have activity \(u_{j2}\) start before the disruption as well, i.e., \(G^1_{u_{j2}} = 0\). The top path yields a length of \(D_{u_{j1},u_{j4} - 2} + D_{u_{j2},u_{j3}} = 0 + 1 = 1\), while the bottom path yields a length of \(D_{u_{j1},u_{j4}} = 2\), which makes the bottom path critical. We can consider the value of \(G_{u_{j1}}\) as the truth assignment of variable \(u_j\). If the variable \(u_j\) is TRUE, the critical path is the top path; if it is FALSE, the critical path is the bottom path.

@@

		\caption{A component corresponding to variable \(u_j,\ j = 1,2,\dots, n\) in the first layer of the constructed activity network for EOIT\_3SAT.}

@@

	\noi The arcs from activities \(u_{j3}\) and \(u_{j4}\) in Figure~\ref{fig:layer1} point to activities in the second layer. Next, we construct the second layer of activities and the arcs connecting the first layer and the second layer, in a similar manner as Figure 2(b) in~\cite{de1997complexity}. To best illustrate the arc construction, we show an example of a clause, \(c_i = u_j \vee u_k \vee \bar{u}_{\ell}\), with two original variables, \(u_j\) and \(u_k\), and one negative complement, \(\bar{u}_{\ell}\). We can consider the activity \(c_{ip}\) corresponding to the truth assignment of the variable \(u_p\), for \(p \in \{j,k,\ell\}\). For the original variables \(u_p\), \(p \in \{j,k\}\), we connect activity \(u_{p4}\) with activity \(c_{ip}\), and activity \(u_{p3}\) with activities \(c_{iq}\), where \(q \in \{j,k,\ell\}, q \neq p\). For the negative complement variable \(u_\ell\), we connect activity \(u_{\ell 3}\) with activity \(c_{i \ell}\), and activity \(u_\ell 4\) with activities \(c_{iq},\ q \in \{j,k\}\). This example by no means limits the clauses to two variables and one negative complement but aims to showcase the procedure of constructing the network components corresponding to a variable and the negative complement of a variable. From now on we refer to the activities representing variables as \(u\)-activities and those representing clauses as \(c\)-activities.\\

@@

	%	& D_{u_{p3},c_{iq}} = 0 \qquad \quad d^1_{u_{p3},c_{iq}} = 0 \quad \forall p \in \{j,k\}, q \in \{j,k,\ell\}, p \neq q\\
	%	& D_{u_{p4},c_{ip}} = 0 \qquad \quad d^1_{u_{p4},c_{ip}} = 0 \quad \forall p \in \{j,k\}\\
	%	& D_{u_{\ell 3},c_{i\ell}} = 0 \qquad \quad d^1_{u_{\ell 3},c_{i \ell}} = 0\\
	%	& D_{u_{\ell 4},c_{iq}} = 0 \qquad \quad d^1_{u_{\ell 4},c_{iq}} = 0 \quad \forall q \in \{j,k\}\\

@@

		\caption{The \(i\)-th clause, \(u_{j} \vee u_{k} \vee \bar{u}_{\ell}\), in the second layer of the constructed activity network for EOIT\_3SAT with the arcs connecting it with the first and the third layer.}

@@

	%Without loss of generality, suppose \(u_j\) is the only TRUE assignment, then activity \(c_{ij}\) can start as early as time \(1\), while the other two \(c\)-activities can only start after time \(2\). 

@@

		\caption{The \(i\)-th clause, \(u_{j} \vee u_{k} \vee u_{\ell}\), in the second layer of the constructed activity network for EOIT\_3SAT with the arcs connecting it with the first and the third layer.}

@@

		Without loss of generality, suppose both \(c_{ij}\) and \(c_{ik}\) can start at time \(1\). This means that both \(u_{j3}\) and \(u_{j4}\) needs to start at time \(1\), which is impossible because no matter what value \(G^1_{u_{j1}}\) takes, at least one activity in \(\{u_{j3},u_{j4}\}\) has to start after time \(2\).

@@

		\(\implies\):  Suppose \(u_j\) is FALSE, then activity \(u_{j4}\) can only start after time \(2\), which will lead to the contradiction that \(c_{ij}\) can start as early as time \(1\). Suppose either \(u_k\) or \(u_\ell\) is TRUE in addition to \(u_j\). Since there is going to be an arc from either activity \(u_{k3}\) or \(u_{\ell 3}\) to activity \(c_{ij}\), and either \(u_{k3}\) or \(u_{\ell_3}\) cannot start before time \(2\), it leads to a contradiction that \(c_{ij}\) can start as early as time \(1\). Therefore, if activity \(c_{ij}\) can start at time \(1\), then \(u_j\) is the only variable which is TRUE.\\
		\(\impliedby\): if only \(u_j\) is TRUE, then \(u_{j4}, u_{k3}\) and \(u_{\ell 3}\) can all start as early as time \(1\), which means that the earliest start time of \(c_{ij}\) is \(1\).

@@

	%For a clause \(i\) that contains variable \(j\), activity \(u_{j4}\) connects to \(c_{ij}\) and activity \(u_{j3}\) connects to the \(c\)-activities which do not correspond to variable \(u_j\). This means that if variable \(u_j\) is assigned a TRUE value, activity \(c_{ij}\) can start as soon as time \(1\) and the other two \(c\)-activities can start no earlier than time \(2\). 

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\renewcommand{\textfraction}{0.0}
\newcommand{\dst}{\displaystyle}
\newcommand{\minx}{\mbox{\( \dst \min_{x \in X} \)}}
\newcommand{\Efx}{\mbox{\( \dst E f (x, \xi) \)}}
\newcommand{\Efxhat}{\mbox{\( \dst E f (\hat{x}, \xi) \)}}
\newcommand{\hxx}{\mbox{\( \hat{x} \)}}
\newcommand{\bpi}{\bar{\pi}}
\newcommand{\xx}{\mbox{\( x \)}}
\newcommand{\txxi}{\mbox{\(\xi\)}}
\newcommand{\var}{\mbox{var}}
\newcommand{\cF}{{\cal F}}
\newcommand{\cA}{{\cal A}}
\newcommand{\cG}{{\cal G}}
\newcommand{\cN}{{\cal N}}
\newcommand{\cO}{{\cal O}}
\newcommand{\txi}{{\xi}}
\newcommand{\PP}{\mbox{\(SP\)}}
\newcommand{\PPn}{\mbox{\(SP_n\)}}
\newcommand{\PPnx}{\mbox{\(SP_{n_x}\)}}
\newcommand{\noi}{\noindent}
\renewcommand{\ss}{\smallskip}
\newcommand{\ms}{\medskip}
\newcommand{\bs}{\bigskip}
\newcommand{\st}{\mbox{s.t.}}
\newcommand{\wpo}{\mbox{wp1}}
\newcommand{\iid}{\mbox{i.i.d.\ }}
\newcommand{\vsmo}{\vspace*{-0.1in}}
\newcommand{\vsmt}{\vspace*{-0.2in}}
\newcommand{\vso}{\vspace*{0.1in}}
\newcommand{\vst}{\vspace*{0.2in}}
\newcommand{\mc}{\multicolumn}
\newcommand{\cP}{{\cal P}}
\newcommand{\tcb}{\textcolor{blue}}
\newcommand{\underv}{\mbox{$\underbar{$v$}$}}
\allowdisplaybreaks

@@

	The management of complex projects through optimization has a rich history in operations research, beginning with the critical path method of \cite{kelley1961criticalpath} (see S{\"o}derlund \cite{soderlund2004building} for an overview). A project is a collection of activities, between which there are precedence relationships due to logical or technological considerations. The precedence relationship is usually reflected as a required separation time between the start time of two activities. Multiple activities can be processed at the same time and there is no limit on how many activities can be processed simultaneously, as long as the precedence requirements are satisfied. See Elmaghraby~\cite{Elmaghraby77} for a detailed treatment of activity networks. In this setting, ``crashing" is an action that consumes a certain amount of resource and shortens the duration of activities accordingly~\cite{kuhl2008dynamic}. The deterministic optimization problem for crashing an activity network was proposed in the 1960s \cite{fulkerson1961network, kelley1961criticalpath}. The goal is to complete all activities in minimum time by allocating resources under one or more budget constraints.\\

@@

@@

	Nomenclature:
	%The notation for the model is displayed as follows:

@@

		\(D_{ik}\)& \(\qquad\) & nominal separation time between activity \(i\) and \(k\), \((i,k) \in \cA\);&\\

@@

		\(B\) & \(\qquad\) & total budget of crashing options;&\\
		\(b_{ij}\) & \(\qquad\) & cost of crashing option \(j\) for activity \(i\), \(i \in I, j \in J_i\);&\\

@@

	\noi We first review the formulation of the deterministic crashing optimization problem in~\cite{fulkerson1961network, kelley1961criticalpath}. A set of activities, \(I\), together with the precedence relationships, \(\mathcal{A} \subseteq I \times I\) form an acyclic activity network \(\mathcal{G} = (I,\mathcal{A})\) which represent the project. An arc \((i,k) \in \mathcal{A}\) indicates that activity \(i\) has to be finished before the start of activity \(k\), and its length, \(D_{ik}\), shows that the start time of activity \(i\) has to be at least \(D_{ik}\) before the start time of activity \(k\). We create two dummy activities \(S, T \in I\) to represent the start and the end of the entire project. Activity \(S\) should precede every activity \(i \in I \backslash \{S\}\) and \(T\) should succeed every activity \(i \in I \backslash \{T\} \), and they both have zero duration, \(D_{Si} = D_{iT} = 0,\ \forall i \in I\). We can apply a finite set of crashing options, \(j \in J_i\), to activity \(i \in I\). One unit application of each option incurs a cost of \(b_{ij}\), while it decreases all required separation lengths starting from activity \(i\) by \(D_{ik}e_{ij},\ \forall (i,k) \in \mathcal{A}\), where \(e_{ij} \in [0,1]\) denotes the unit effectiveness of the crashing option \(j\). For example, suppose the required separation between the start time of activity \(1\) and \(2\) is \(D_{12} = 10\), and applying one unit of crashing option \(1\) on activity \(1\) decreases that to its half; i.e., \(e_{11} = 0.5\). If we apply \(0.4\) unit of crashing option \(1\) on activity \(1\), \(x_{11} = 0.4\), the required separation between activity \(1\) and \(2\) becomes \(10(1 - 0.4 \times 0.5) = 8\). The total cost of crashing cannot exceed a given budget, \(B\). Given $t_S \equiv 0$, the objective is then to minimize the start time of activity \(T\). Thus, we formulate the deterministic project crashing problem as:
	For a project crashing problem under stochastic disruptions, we assume at most one stochastic disruption can occur at a random time in the project span, since for the applications we are interested in (natural disasters, terrorist attacks, major market crashes, etc.), it is unlikely for two disruptions to occur in entire the time horizon. For example, suppose we manage a construction project, and we aim to plan against the potential hazard caused by earthquakes. It is unlikely for two major earthquakes to hit the same area twice within a year (aftershocks can be considered part of a major earthquake.) It is also reasonable to assume for each activity \(i \in I\), the crashing decision needs to be made at the start of that activity, and cannot be changed afterward, especially when contracts are involved in commitment of resources~\cite{oberlender1993project}. The disruption does not affect the started activities but changes the length of those that have not started according to a random distribution. It is usually hard to compute the recourse function directly when parameters are distributed according to continuous distributions, and therefore SAA is proposed to approximate the problem by a large-scale deterministic optimization problem with a finite scenario set \cite{kim2015guide,shapiro2009lectures}. In this paper, we assume there is a discrete finite set of scenarios indexed by \(\omega \in \Omega\). For each scenario \(\omega\), the random realization of parameters, \(\xi^\omega\), consists of the timing of disruption, \(H\), and the changed duration between activity \(i\) and \(k\), \(d_{ik}, \forall (i,k) \in \cA\). \\
	The extensive formulation of the two-stage stochastic program is shown as formulation~\eqref{prob:extensive}:

@@

		\text{s.t.} \quad & t_k - t_i \geq D_{ik}(1 - \sum_{j \in J_i} x_{ij} e_{ij}) \qquad \qquad \forall \, (i,k) \in \mathcal{A} \label{cons:Sep}\\

@@

		& t_k^\omega - t_i^\omega \geq D_{ik} + d_{ik}^\omega G_i^\omega -\sum_{j \in J_i} D_{ik} e_{ij} x_{ij}^\omega - \sum_{j \in J_i} d_{ik}^\omega e_{ij} z_{ij}^\omega \qquad \qquad \forall \,(i,k) \in \cA, \omega \in \Omega \label{cons:scenSep}\\

@@

	In model~\eqref{prob:extensive}, we are minimizing the expected project span by taking the sum of each scenario's project span weighted by the scenario probability. We keep the constraints~\eqref{cons:dSep}-\eqref{cons:dSingleBudget} for the nominal scenario as~\eqref{cons:Sep}-\eqref{cons:SingleBudget}. In constraints~\eqref{cons:G1}-\eqref{cons:G2}, variables \(G^\omega_i\) takes value \(1\) if activity \(i\) starts after the disruption time; otherwise it takes value 0, and \(M\) is a large number to enforce this logic relationships. This is important in our problem setting because the duration of each activity depends on its temporal relationship to the disruption time, which is reflected in constraint~\eqref{cons:scenSep}. Also we have to ensure that the decisions made before the disruption time in each scenario are the same as the nominal decisions. Therefore, we set up non-anticipativity constraints~\eqref{cons:tG1}-\eqref{cons:xG2}. For each scenario, the duration between activity \(i\) and \(k\) becomes \((D_{ik} + d_{ik}^\omega G_i^\omega)(1 - \sum_{j \in J_i} e_{ij}x_{ij}^\omega)\), which expands to the form of constraint~\eqref{cons:scenSep}. If \(G_i^\omega = 0\), which means activity \(i\) starts before the disruption time of scenario \(\omega\), this expression is the same as \(D_{ik} (1 - \sum_{j \in J_i} e_{ij}x_{ij})\) since \(x_{ij} = x_{ij}^\omega\) is enforced by constraints~\eqref{cons:xG1} and~\eqref{cons:xG2}. If \(G_i^\omega = 1\), then the duration between activity \(i\) and \(k\) is changed to \(D_{ik} + d_{ik}^\omega\). The expression \((D_{ik} + d_{ik}^\omega G_i^\omega)(1 - \sum_{j \in J_i} e_{ij}x_{ij}^\omega)\) contains a bilinear term \(G_i^\omega x_{ij}^\omega\), which can be linearized by introducing a binary variable \(z_{ij}^\omega\) and constraints \eqref{cons:linearize1}-\eqref{cons:linearize3}.

@@

	In this section, we prove that the project crashing optimization problem under a stochastic disruption is NP-hard. The proof process is largely based on the proof of De et al.~1997 for the discrete time-cost tradeoff (\verb|DTCT|) problem. In De et al.~\cite{de1997complexity}, the authors show that an exactly-one-in-three \verb|3SAT| (\verb|EOIT_3SAT|) problem can be transformed to an instance of the \verb|DTCT| decision problem in polynomial time. Since \verb|EOIT_3SAT| is NP-complete (see \cite{Garey1979ComputersAI} for the detailed proof), the authors prove that the \verb|DTCT| decision problem is also NP-complete, and the \verb|DTCT| optimization problem is NP-hard.\\

@@

	Similar to De et al.~\cite{de1997complexity}, we also first introduce the \verb|EOIT_3SAT| problem: let \(U = \{u_1,u_2, \dots, u_n\}\) be a set of variables. A literal can be either \(u\) or \(\bar{u} = \neg u\) for a \(u \in U\). Let \(C = \{c_1, c_2, \dots, c_m\}\) be a set of clauses, each of which is formed by a disjunction of three literals. The \verb|EOIT_3SAT| problem asks for whether there is a truth assignment for \(U\) such that each clause in \(C\) has exactly one true literal. \\

@@

	Based on the above instance of \verb|EOIT_3SAT|, we formulate an instance of the activity network for the project crashing problem. The activity network consists of three layers of nodes: the first layer contains \(3n\) nodes and represents the truth assignment of each variable; the second layer contains \(3m\) nodes and represents the value of clauses; the third layer contains node \(T\) which represents the end of the project. The first layer has \(n\) components in parallel, each of which corresponds to a variable and contains 3 nodes connected, denoted as \(u_{j1},\ u_{j2}\) and \(u_{j3}\) for the component \(j\), as shown in Figure~\ref{fig:layer1}. Figure~\ref{fig:layer1} also displays how the first layer connects to the third layer. We set the scenario set as a singleton \(\Omega = \{1\}\) and set the disruption time as \(H^1 = 0\) with probability \(p^1 = 1\). We set parameter values associated with the arcs in Figure~\ref{fig:layer1} as:

@@

	& D_{u_{j1},u_{j2}} = 1 \qquad \quad d^1_{u_{j1},u_{j2}} = 1\\
	& D_{u_{j1},u_{j3}} = 2 \qquad \quad d^1_{u_{j1},u_{j3}} = -1  \\
	& D_{u_{j3},T} = 0 \qquad \quad \;\; d^1_{u_{j3},T} = 0.

@@

	We do not allow any activities within the first layer to be crashed, i.e., \(J_{u_{jk}} = \emptyset\), for all \(j = 1,2,\dots, n,\ k = 1,2,3\). With this setup, the length of the critical path for the \(j\)-th component is always \(2\) for all \(j = 1,2,\dots,n\), as it is optimal to have activity \(u_{j1}\) start at time \(0\). However, whether the critical path traverses activities \(u_{j2}\) (top path) or the activity \(u_{j3}\) (bottom path) depends on the value of \(G^1_{u_{j1}}\). If \(G^1_{u_{j1}} = 1\), the \(G\) variables will take value \(1\) for all subsequent nodes. The top path yields a length of \(D_{u_{j1},u_{j2}} + d^1_{u_{j1},u_{j2}} = 1 + 1 = 2\), while the bottom path yields a length of \(D_{u_{j1},u_{j3}} + d^1_{u_{j1},u_{j3}} = 2 - 1 = 1\), which makes the top path critical. If \(G^1_{u_{j1}} = 0\), it is optimal to have activity \(u_{j2}\) start before the disruption as well, i.e., \(G^1_{u_{j2}} = 0\). The top path yields a length of \(D_{u_{j1},u_{j2}} = 1\), while the bottom path yields a length of \(D_{u_{j1},u_{j3}} = 2\), which makes the bottom path critical. We can consider the value of \(G^1_{u_{j1}}\) as the truth assignment of variable \(u_j\). If the variable \(u_j\) is TRUE, the critical path is the top path; if it is FALSE, the critical path is the bottom path. 

@@

		\caption{A component corresponding to variable \(u_j,\ j = 1,2,\dots, n\) in the first layer of the constructed activity network for \texttt{EOIT\char`_3SAT}.}

@@

	\noi The arcs from activities \(u_{j2}\) and \(u_{j3}\) in Figure~\ref{fig:layer1} point to activities in the second layer. Next, we construct the second layer of activities and the arcs connecting the first layer and the second layer, in a similar manner as Figure 2(b) in~\cite{de1997complexity}. To best illustrate the arc construction, we show an example of a clause, \(c_i = u_j \vee u_k \vee \bar{u}_{\ell}\), with two original variables, \(u_j\) and \(u_k\), and one negative complement, \(\bar{u}_{\ell}\). We can consider the activity \(c_{ip}\) corresponding to the truth assignment of the variable \(u_p\), for \(p \in \{j,k,\ell\}\). For the original variables \(u_p\), \(p \in \{j,k\}\), we connect activity \(u_{p3}\) with activity \(c_{ip}\), and activity \(u_{p2}\) with activities \(c_{iq}\), where \(q \in \{j,k,\ell\}, q \neq p\). For the negative complement variable \(u_\ell\), we connect activity \(u_{\ell 2}\) with activity \(c_{i \ell}\), and activity \(u_{\ell 3}\) with activities \(c_{iq},\ q \in \{j,k\}\). This example by no means limits the clauses to two variables and one negative complement but aims to showcase the procedure of constructing the network components corresponding to a variable and the negative complement of a variable. From now on we refer to the activities representing variables as \(u\)-activities and those representing clauses as \(c\)-activities.\\

@@

@@

		\caption{The \(i\)-th clause, \(u_{j} \vee u_{k} \vee \bar{u}_{\ell}\), in the second layer of the constructed activity network for \texttt{EOIT\char`_3SAT} with the arcs connecting it with the first and the third layer.}

@@

@@

		\caption{The \(i\)-th clause, \(u_{j} \vee u_{k} \vee u_{\ell}\), in the second layer of the constructed activity network for \texttt{EOIT\char`_3SAT} with the arcs connecting it with the first and the third layer.}

@@

		Without loss of generality, suppose both \(c_{ij}\) and \(c_{ik}\) can start at time \(1\). This means that both \(u_{j2}\) and \(u_{j3}\) needs to start at time \(1\), which is impossible because no matter what value \(G^1_{u_{j1}}\) takes, at least one activity in \(\{u_{j2},u_{j3}\}\) has to start after time \(2\).

@@

		\(\implies\):  Suppose \(u_j\) is FALSE, then activity \(u_{j3}\) can only start after time \(2\), which will lead to the contradiction that \(c_{ij}\) can start as early as time \(1\). Suppose either \(u_k\) or \(u_\ell\) is TRUE in addition to \(u_j\). Since there is going to be an arc from either activity \(u_{k2}\) or \(u_{\ell 2}\) to activity \(c_{ij}\), and either \(u_{k2}\) or \(u_{\ell 2}\) cannot start before time \(2\), it leads to a contradiction that \(c_{ij}\) can start as early as time \(1\). Therefore, if activity \(c_{ij}\) can start at time \(1\), then \(u_j\) is the only variable which is TRUE.\\
		\(\impliedby\): if only \(u_j\) is TRUE, then \(u_{j3}, u_{k2}\) and \(u_{\ell 2}\) can all start as early as time \(1\), which means that the earliest start time of \(c_{ij}\) is \(1\).

@@

	\noi A direct result from Lemma~\ref{lemma:iff} shows if there exists a truth assignment that meets the requirement of \verb|EOIT_3SAT|, there are at most \(2m\) number of \(c\)-activities (2 per clause) that has to start after time \(2\). Since we have a total of \(B = 2m\) units of resource, we can crash all of those \(c\)-activities to achieve a project length of \(2\). This means that we can transform an \verb|EOIT_3SAT| problem to a project crashing decision problem under a stochastic disruption using the activity network construction process above. If we can find a feasible \verb|EOIT_3SAT| assignment, it is possible to obtain a project length less than or equal to 2. Next, we show the formal NP-hardness proof of the project crashing problem with a stochastic disruption. We first formally define the project crashing decision problem under a stochastic disruption:
